Homicide Increases in the COVID-19 Era

### 2025-03-04 Draft2.qmd
organization

@@

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) + 
  ggtitle("Correlations Between the Number of Crimes and \n Different Law Enforcement and Socioeconomic Factors")
```

## geom_jitter
## geom_jitter

```{r, echo=FALSE}
ggplot(crime_la, aes(x = DateParsed, y = Vict.Sex, color = Vict.Sex)) +
  geom_jitter(alpha = 0.4, width = 0.2) +
  labs(title = "Crime Over Time by Gender",
       x = "Date",
       y = "Gender") +
  theme_minimal()

## Density Plot

crime_la$Hour <- as.numeric(format(crime_la$DateParsed, "%H"))

ggplot(crime_la, aes(x = Hour, fill = ..density..)) +
  geom_density(alpha = 0.5, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Crime Density Per Hour",
       x = "Hour",
       y = "Density") +
  theme_minimal()

@@

```{r, echo=FALSE}
# Data Preparation
homicides <- crime_la %>% inner_join(crimes_with_zips, by = "crime_id") %>% filter(offense_type == "HOMICIDE") %>% mutate(date = as.Date(date)) 
homicides <- homicides %>% left_join(cali_median_income, by = c("zip_code" = "ZCTA"))
homicides$median_income <- as.numeric(homicides$median_income)

# Form income quartiles
homicides <- homicides %>% mutate(income_quartile = ntile(median_income, 4)) 
homicide_trends <- homicides %>% mutate(year = lubridate::year(date)) %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), .groups = "drop")

# Plot homicide trends over time by income quartile
ggplot(homicide_trends, aes(x = year, y = homicide_count, color = factor(income_quartile))) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    x = "Year",
    y = "Number of Homicides",
    color = "Income Quartile"
  ) +
  theme_minimal()
```

## Plot 6

@@

>>>>

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "..")
```

@@

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) +
  labs(fill = "Correlation")
```
### Interpretation
This heat map shows the correlations between different variables that could affect the number of crimes committed. Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles. The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.

Here, we see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions. Years is positively correlated with number of crimes, meaning that as time passed, the number of crimes increased. Similarly, as the number of evictions increases, so does the number of crimes. Interestingly, as the unemployment rate increases, the number of crimes decreases, which is the opposite effect of what was expected (one would anticipate that higher unemployment would result in more crime). The LAPD budget interestingly had a medium-strength positive correlation with number of crimes, suggesting we should further examine the increase in the LAPD budget compared to the increase in crimes.

## Homicide Victims by Age and Gender in Los Angeles 2010-Present

```{r, echo=FALSE}
homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))

levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
  geom_jitter(alpha = 0.4, width = 0.2) +
  labs(x = "Gender",
       y = "Age") +
  theme_minimal() +
  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
  theme(legend.position = 'none')
```
### Interpretation

This jitter plot shows us both the frequencies and the demographics of homicide victims in Los Angeles from 2010 to present. Based on the density of the points, we see that most homicide victims in the last 15 years have been men between the ages of around 15 to 65. Additionally, we see that there are not many victims who are toddlers and younger children, but there is a good number of babies who have been killed. This is an intricacy in our data that might be interesting to look into further, particularly into the circumstances of those deaths and potential correlations between reports of child abuse and/or domestic violence and infanticide.

## Frequency of Crimes in Los Angeles Throughout the Day

```{r, echo=FALSE}
crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))

ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(x = "Hour (Military Time)",
       y = "Density") +
  theme_minimal() +
  theme(legend.position = 'none')
```
### Interpretation

This density plot shows the frequency of crimes depending on the hour (in military time) that they occurred. Based on the plot, we see that fewer crimes occur early in the morning (with the lowest at 5 AM). The data peaks at noon, meaning the most crimes occur in that hour of the day. Crimes are more frequent after that, until the early morning. The second most common crime occurrence is 6 PM, or 18 hours.

@@

```{r, echo=FALSE}
# Data Preparation
homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
  mutate(date = as.Date(DateParsed),
         zip = as.character(ZCTA5CE20)) %>% select(-X)

homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)

# Form income quartiles
homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
          ref = mean(S1903_C03_001E),
          .groups = "drop")

homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
# removing NAs
homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))

# creating labels
levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
# Plot homicide trends over time by income quartile
ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
  labs(
    x = "Year",
    y = "Number of Homicides",
    fill = "Income Quartile",
    title = "Homicide Rates in Los Angeles By Median Income Level"
  ) +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma", direction = -1)
```

### Interpretation
Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic. In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic. On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.

@@

# Spatio-Temporal Analysis of Crime in Los Angeles

```{r}
crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv")
